---
title: "Homework 11"
author: "BMV"
date: "4/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Question 1:**

I have 1211 red spruce individuals sampled across the elevational gradient on Camels Hump Mountain. For each individudal, I have the diameter at breast height (DBH) (lower limit of 2cm). I would like to use this measurement to estimate the basal area for each tree. I am using the formula below that converts DBH measured in cm to basal area in m^2.

Tree Basal Area (TBA) = (DBH/200)^2 x 3.142 m2.

Where DBH is the Diameter at Breast height (cm) and 3.142 is π

Call in the data for global variables. For this data set, I do not have a parameter that can vary.
```{r}
data=read.csv("/Volumes/Macintosh HD/Users/Brittany/Documents/UVM/red spruce/CHSurvey1965_2015/CH_data_1965_2015.csv", header=TRUE)

picea=subset(data, data$SPECIES==4)

picea=picea$DIAM
head(picea)
```

```{r}
##################################################
# function: TreeBasalArea
# converts the diameter in cm to the basal area in m2
# input: DBH=vector of diameter of breast heights for picea
# output: TBA is a vector of basal area estimates in meters sqaured
#------------------------------------------------- 
TreeBasalArea=function(DBH=picea){
  TBA=(DBH/200)^2 * 3.142
return(TBA)
}

head(TreeBasalArea())

##################################################
# function: BasalAreaPlot
# plot basal area 
# input: DBH=vector of diameter of breast heights for picea
# output: smoothed curve with parameters in graph
#------------------------------------------------- 
BasalAreaPlot <- function(DBH=picea) {
plot(x=DBH,y=TreeBasalArea(DBH),xlab="Diameter at  breast height (DBH)",ylab="Basal area")
      return()
}
BasalAreaPlot()
```

**Question 2:**

Using a different data set that contains estimates of timing of bud set for 42 trees located along elevational gradient. Looking at relationship between bud set ~ elevation. 

Preliminaries
```{r}
library("ggplot2")
library("TeachingDemos")
char2seed("Cruel April")
```

Call in data to global environment.
```{r}
budset=read.csv("/Volumes/kellrlab/datashare/Spruce/Collection Data/Seeds/GerminationRound1/budset/ele_mean_budset.csv", header=TRUE)

budset=budset[,c(2,4)]
head(budset)
```

Functions
```{r}

###################################
# Function: readData
# read in or generate data frame
# input: file name (or nothing for demo)
# output: 3-column data frame of observed data (ID, x, y)
#---------------------------------
readData <- function(z=NULL) {
  if (is.null(z)) {
    xVar <- budset$Elevation
    yVar <- budset$Julian_Day
    dF <- data.frame(ID=seq_along(xVar),xVar,yVar)
  }
  return(dF)
}
readData()

###################################
# Function: getMetric
# calculate metric for randomization test
# input: 3-column data frame for regression
# output: regression slope
#---------------------------------
getMetric <- function(z=NULL) {
  if (is.null(z)) {
    xVar <- budset$Elevation
    yVar <- budset$Julian_Day
    z <- data.frame(ID=seq_along(xVar),xVar,yVar)
  }
  . <- lm(z[,3]~z[,2])
  . <- summary(.)
  . <- .$coefficients[2,1]
  slope <- .
  return(slope)
}
getMetric()

###################################
# Function: shuffleData
# randomize data for regression analysis
# input: 3-column data frame (ID, xVar, yVar)
# output: 3-column data frame (ID, xVar, yVar) 
#---------------------------------
shuffleData <- function(z=NULL) {
  if (is.null(z)) {
    xVar <- budset$Elevation
    yVar <- budset$Julian_Day
    z <- data.frame(ID=seq_along(xVar),xVar,yVar)
  }
  z[,3] <- sample(z[,3])
  return(z)
}
shuffleData()

###################################
# Function: getPVal
# calculate p value for observed, simulated data
# input: list of observed metric and vector of simulated metric
# output: lower, upper tail probability vector
#---------------------------------
getPVal <- function(z=NULL) {
  if (is.null(z)) {
    z <- list(xObs=runif(1),xSim=runif(1000))
  }
  pLower <- mean(z[[2]] <= z[[1]])
  pUpper <- mean(z[[2]] >= z[[1]])
  return(c(pL=pLower,pU=pUpper))
}
getPVal()

###################################
# Function: plotRanTest
# ggplot graph
# input: list of observed metric and vector of simulated metric
# output: ggplot graph
#---------------------------------
plotRanTest <- function(z=NULL) {
  if (is.null(z)) {
    z <- list(xObs=runif(1),xSim=runif(1000))
  }
  dF <- data.frame(ID=seq_along(z[[2]]),simX=z[[2]])
  p1 <- ggplot(data=dF, mapping = aes(x=simX))
  p1 + geom_histogram(mapping = aes(fill=I("lightsalmon"),color=I("black"))) + geom_vline(aes(xintercept=z[[1]],col="blue"))
}
plotRanTest()
```

Main body of code
```{r}
nSim <- 1000 # number of simulations
XSim <- rep(NA, nSim) # will hold simulated slopes

dF <- readData()
Xobs <- getMetric(dF)

for (i in seq_len(nSim)) {
  XSim[i] <- getMetric(shuffleData(dF))
}

slopes <- list(Xobs,XSim)
getPVal(slopes)
plotRanTest(slopes)
```

**Question 3**

The p-values are nearly the same in both tests. I would expect a slight difference given the randomization.  

```{r}
lmBudset <- lm(budset$Julian_Day ~ budset$Elevation, data = readData())
summary(lmBudset)


lmBudsetShuffle <- lm(budset$Julian_Day ~ budset$Elevation, data = shuffleData())
summary(lmBudsetShuffle)
```

