---
title: "Homework 12"
author: "Brittany Verrico"
date: "4/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Generate a fake data set to make figures for local adaptation hypotheses. The reaction norm will be continuous because I am utilizing the transfer distance approach. 

```{}
Transfer distance = (Climate at collection site) - (Climate at planting site)
```

If populations are locally adapted, then fitness is highest when the transfer distance is 0. That is, the home and planting site climatic conditions (based on a particular variable) are the same/similar. Fitness should decrease as transfer distance increases. 

I have 3 common gardens spaced along an elevational gradient, so I will demonstrate what I expect for local adaptation at each garden. I will use mean annual temperature as my climate variable, and I have already figured out that my transfer distances fall between 0 and 4 degrees C. Will not go into all the details here. I would like to demonstrate what no local adaptation looks like in each figure. 

I would also like to plot the range of transfer distances for collections of trees collected at two scales. Again, not going into the details. 

Load in the required packages
```{r}
library(ggplot2)
library(patchwork)
library(cowplot)
```

Generate a fake, but realistic data set. 

Low elevation
```{r}
lowSeq=seq(0,4,.01)
# create a normal distribution, with a mean of 0
lowdensities<-dnorm(lowSeq, mean=0,sd=0.6)
# but only keep those value greater than 0 because I will not have negative td in my experiment
lowdensities2=lowdensities[lowdensities>=0]
BroadScale=rep("Broad-scale", length(lowdensities2))
# put everything in a data frame
lowMat=data.frame(x=lowSeq, y=lowdensities2, scale=BroadScale)
```

No local adaptation at low elevation
```{r}
# create a fake y value
noLALow=rep(0.2,length(lowdensities2))
FineScale=rep("Fine-scale", length(lowdensities2))
noLAmat=data.frame(x=lowSeq, y=noLALow, scale=FineScale)

# create combined matrix for plotting
plot_matLow=rbind(lowMat, noLAmat)
```

```{r}
# plot local adaptation of low elevation
Low=ggplot(data= plot_matLow, mapping = aes(x=x, y=y, group=factor(scale))) + geom_line(aes(color=scale), size=1) + theme_classic() + 
  ylab("Fitness") +
  xlab("Transfer Distance (°C)")+ 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_text(size=12, face="bold"),
        axis.title.y = element_text(size=12, face="bold"),
        axis.text.x = element_text(face="bold",size=11), legend.position = "none") + scale_color_manual(values=c('red','black')) +
  geom_text(aes(3,1, label = "Low Garden"), size=4)
print(Low)
```

Mid elevation
```{r}
mid<-seq(-2,2,.01)
middensities<-dnorm(mid, mean=0,sd=0.3)
BroadScale=rep("Broad-scale", length(lowdensities2))
midMat=data.frame(x=mid, y=middensities, scale=BroadScale)
```

No local adaptation at mid elevation
```{r}
# create a fake y value
noLAmid=rep(0.2,length(lowdensities2))
FineScale=rep("Fine-scale", length(middensities))
noLAmatMid=data.frame(x=mid, y=noLAmid, scale=FineScale)

# create combined matrix for plotting
plot_matMid=rbind(midMat, noLAmatMid)
```

```{r}
# plot local adaptation of low elevation
Mid=ggplot(data= plot_matMid, mapping = aes(x=x, y=y, group=factor(scale))) + geom_line(aes(color=scale), size=1) + theme_classic() + 
  ylab("Fitness") +
  xlab("")+ 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_text(size=12, face="bold"),
        axis.title.y = element_text(size=12, face="bold"),
        axis.text.x = element_text(face="bold",size=11), legend.position = "none") + scale_color_manual(values=c('red','black')) +
  geom_text(aes(1,2.25, label = "Mid Garden"), size=4)
print(Mid)
```

High elevation
```{r}
high<-seq(-4,0,.01)
highdensities<-dnorm(high, mean=0,sd=0.3)
highdensities2=highdensities[highdensities>=0]
BroadScale=rep("Broad-scale", length(highdensities2))
highMat=data.frame(x=high, y=highdensities2, scale=BroadScale)
```

No local adaptation at high elevation
```{r}
# create a fake y value
noLAhigh=rep(0.2,length(highdensities2))
FineScale=rep("Fine-scale", length(highdensities2))
noLAmatHigh=data.frame(x=high, y=noLAhigh, scale=FineScale)

# create combined matrix for plotting
plot_mat=rbind(highMat, noLAmatHigh)
```

```{r}
# plot local adaptation of low elevation
High=ggplot(data= plot_mat, mapping = aes(x=x, y=y, group=factor(scale))) + geom_line(aes(color=scale), size=1) + theme_classic() +
  ylab("Fitness") +
  xlab("")+ 
  theme(axis.text.y = element_blank(), axis.ticks = element_blank(), axis.title.x = element_text(size=12, face="bold"),
        axis.title.y = element_text(size=12, face="bold"),
        axis.text.x = element_text(face="bold",size=11), legend.position = "top") + scale_color_manual(values=c('red','black')) +
  geom_text(aes(-1,2.25, label = "High Garden"), size=4)
print(High)
```


If patchwork does not work, then this can be an alternative. 
```{r}
plot_grid(Low, Mid, High, labels = c("Low", "Mid", "High"), align = "v", nrow=3) + theme(plot.margin = unit(c(0,6,0,6), "cm"))
#ggsave(filename = "/Volumes/kellrlab/brittany/dissertation/Grants/Marvin/LocalAdaptationHo.pdf", plot=plotMe, device = "pdf")
```


Plot transfer distances between scales.
```{r}
td=read.csv("/Volumes/kellrlab/datashare/Spruce/MansfieldCG/CommonGardenPlans/TDplot.csv", header=TRUE)
head(td)

tdplot=ggplot(data=td, aes(x=td, y=garden)) + geom_point(aes(shape =scale), size = 2) + geom_point(aes(color =tally)) +
  scale_x_continuous(breaks=seq(from=-4, to=4, by=1))+ theme_bw() +
  theme(panel.grid = element_blank(), axis.text.y = element_blank(), axis.title.y = element_blank(), axis.text.x = element_text(face="bold"), axis.title.x = element_text(face="bold"), legend.position = "top") + labs(x="Transfer Distance (°C)")+ geom_hline(yintercept = 2.5) + 
   geom_hline(yintercept = 4.5) +
geom_text(aes(3,2.25, label = "Low Garden"), size=4) +
geom_text(aes(3,4.25, label = "Mid Garden"), size=4) +
  geom_text(aes(3,6, label = "High Garden"), size=4) +
  scale_colour_gradient(low = "red",high = "blue") +
  guides(colour=FALSE)
print(tdplot)
```

Try plotting everything together.
This could look better, but I think it looks okay for now.
```{r}
High + Mid + Low + plot_layout(ncol=1) | tdplot
```