---
title: "Homework 8"
author: "BMV"
date: "3/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description

I am going to create a fake dataset that creates a relationship between budset and elevation. In my experiment, I expect phenotypic variation in ecologically important traits related to growing season length. I hypothesized that bud set occurs earlier at higher elevations and later at lower elevations. A preliminary analysis indicated that the difference may be on the scale of 1 week, with a negative relationship between time of bud set and elevation. 

I could use a regression or an ANOVA to analyze the data, depending if I define the independent variable (elevation) as continuous or discrete. To use elevation as a discrete variable, I will need to bin trees into  elevational groups (low, mid, high). For this assignment, I will use an ANOVA.

I have a sample size of 40 trees in my experiment. I will split the families into three groups (n=13). The timing of budset should follow a normal distribution (per assignment directions). Generally, I will have ~ 1 week difference in timing of budset between low and high elevations. The mid-elevation timing could be anywhere between low and high. The standard deviation is a measure of days. Perhaps the high elevation group will have a smaller sd, because of how different the climate and environment is at the summit compared to the low and mid elevations. 

```{r}
# Preliminaries
library(reshape2)
library(ggplot2)

# Global objects

##### first for loop
# define an empty vector to hold the p values
results=NULL
# define the number of times to iterate the for loop
times=1000

###### second for loop
# establish possible means for budset 
# i am setting this to a maximum difference of 10 days across the elevation
possibleMeans=seq(280,290, by=1)

# set the number of combinations
# i am setting this to three because i have 3 elevations
y=3

# find all combinations with this nice package
all_combinations <- combn(possibleMeans,y)

# create a new object with transpose function
# change from wide to long format
# this is needed for easy use in the for loop
meanCombos=t(all_combinations)

# create empty vector for the for loop
results2=NULL
product=NULL
# create empty matrix for the for loop 
# number of rows equals number of combinations
myMat=matrix(nrow=165, ncol=4)


############################################################
# FUNCTION: dataCreate
# Create a fake data set that establishes a random distribution for timing of budset/elevation following a normal distribution
# input: 3 means and 3 standard deviations corresponding to low, mid, and high elevation, a value of n (sample size)
# output: a data frame with 3 columns representing the three elevations. each column will have 13 rows representing a random sample of 13 points from the normal distribution.
# ----------------------------------------------------------
dataCreate <- function(mean1=280,mean2=283,mean3=287,n=13,sd1=4,sd2=3,sd3=2){
  mydf <- data.frame(low=rnorm(mean=mean1,n=n,sd=sd1), 
                     mid=rnorm(mean=mean2,n=n,sd=sd2), 
                     high=rnorm(mean=mean3,n=n,sd=sd3))
  return(mydf)
}
dataCreate()
# assign the output from the function to a variable
myDF <- dataCreate()
############################################################
############################################################
# FUNCTION: ANOfun
# Transform the data from wide to long format with reshape package. Perform the ANOVA analysis.
# input: the data frame created from the first function (myDF) that holds the fake data, and a second dataframe which is the transformed version of the first data frame
# output: a data frame with the summary output of the ANOVA (ANOsum)
# ----------------------------------------------------------
ANOfun=function(data1=myDF){
  suppressMessages(expr = myDFlong <- melt(data1))
  # assign names for easy reference in the model
  colnames(myDFlong)=c("elevation", "budset")
  # create an object that will hold the output from the ANOVA
  ANOmod=aov(budset ~ elevation, data=myDFlong)
  # obtain the summary output from the ANOVA
  ANOsum=summary(ANOmod)
  return(ANOsum)
}
ANOfun() 
############################################################
############################################################
# FUNCTION: PLOTfun
# plot the relationship between time of budset and elevation
# input: fake data set in the wide format
# output: figure showing the relationship
# ----------------------------------------------------------
PLOTfun=function(data=myDF){
  # re-create the wide format of the data set. i only called this in the previous function.
  myDFlong <- melt(data)
  # assign names for easy reference in the model
  colnames(myDFlong)=c("elevation", "budset")
  ggplot(data=myDFlong, aes(x=elevation, y=budset, fill=elevation)) + geom_boxplot() + theme_bw()
}
PLOTfun()
############################################################
############################################################
# Program body

# Run the analysis several times to see the number of times p is <= 0.05.
for(i in 1:times){
  myDF <- dataCreate()
  ANOsum <- ANOfun()
  # create a vector that holds the p-value from the ANOVA summary
  ANOfunpval <- ANOsum[[1]][["Pr(>F)"]][1]
  # place the vector of 1 into the results vector, that will continously have an additional element (totally the number of times at the end of the loop)
  results[i]=ANOfunpval
}
sum(results<=0.05) 


# determine the possible combinations of means that would lead to significant differences in budset across the elevation
for(j in 1: nrow(meanCombos)){
for(i in 1:50){
  myDF <- dataCreate(mean1 = meanCombos[j,1], mean2= meanCombos[j,2], mean3 = meanCombos[j,3])
  ANOsum <- ANOfun()
  # create a vector that holds the p-value from the ANOVA summary
  ANOfunpval <- ANOsum[[1]][["Pr(>F)"]][1]
  results2[i]=ANOfunpval
  product[i]=(sum(results2<=0.05))
  # add in mean values for each elevation into the matrix
  myMat[j,1]=meanCombos[j,1]
  myMat[j,2]=meanCombos[j,2]
  myMat[j,3]=meanCombos[j,3]
  myMat[j,4]=product[i]
}
}
print(myMat)
```

Code that I could not get to work:
```
############################################################
# FUNCTION: COMBOfun
# determine the number of significant p values based on several combinations of mean values for each elevation
# input: dataframe including all possible mean values
# output: matrix with the combinations listed and the number of significant results
# ----------------------------------------------------------
COMBOfun=function(comboData=meanCombos){

results2=NULL
product=NULL
myMat=matrix(nrow=165, ncol=4)

  for(j in 1: nrow(meanCombos)){
  for(i in 1:50){
  myDF <- dataCreate(mean1 = comboData[j,1], mean2= comboData[j,2], mean3 = comboData[j,3])
  ANOsum <- ANOfun()
  # create a vector that holds the p-value from the ANOVA summary
  ANOfunpval <- ANOsum[[1]][["Pr(>F)"]][1]
  results2[i]=ANOfunpval
  product[i]=(sum(results2<=0.05))
  # add in mean values for each elevation into the matrix
  myMat[j,1]=comboData[j,1]
  myMat[j,2]=comboData[j,2]
  myMat[j,3]=comboData[j,3]
  myMat[j,4]=product[i]
  return(myMat)
}
}
}
COMBOfun()
############################################################
```